---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
```
# chaint
## "chain with a tee" - tee functions with magrittr/dplyr chains

## Installation:

```
devtools::install_github("jonocarroll/chaint")
```

## Motivation:

`dplyr` chains are undoubtedly more readable than composed functions, and this makes processing of sequential steps much simpler. When written with line breaks, such as 

```{r}
mtcars %>% 
   group_by(cyl) %>% 
   summarise(abc = mean(mpg))
```

individual processing steps can be skipped over by commenting out a single line

```{r}
mtcars %>% 
#   group_by(cyl) %>% 
   summarise(abc = mean(mpg))
```

The ouptut can then be viewed, and processing resumed.

Often we are interested in what is happening in the middle of this chain, or would like to do something conditionally. `magrittr::%T>%` provides a tee-like function which essentially boils down to

```{r, eval=FALSE}
%T>% <- function(lhs, rhs) {
  eval(rhs)
  return(lhs)
}
```
This is useful for passing data to the `rhs` function which has side effects and for which the return value is to be discarded. However, `%T>%` doesn't take arguments, and the `rhs` function is required to do something with the incomming data.

`chaint` provides a similar function `tee` which is a more sophisticated version of `%T>%` which

    - still takes the data as input to the first argument, but
    - doesn't require the data to be used in any way, 
    - provides a mechanism for adding messages within a chain,
    - can be conditional on the data,
    - can take arguments to be passed on to the `rhs` function.

## Example usage:

```{r}
library(chaint)
mtcars %>% 
   group_by(cyl) %>% 
   say("caution, grouped data!") %>% 
   summarise(abc = mean(mpg))
```
